Exploring Molecular Conformations and Related Time Scales via Nonequilibrium Molecular Dynamics Simulation

### 2013-08-09 writedown/npt-neq/alanine.tex
added discussion of the dipole moment and orientation

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{fig.ala/ext.mode1.100.Ex.01.00.t1200ps/fig-begin-1.eps}
  \includegraphics[width=0.45\textwidth]{fig.ala/ext.mode1.100.Ex.01.00.t1200ps/fig-end-1.eps}
  \caption{The population of metastable states of the alanine dipeptide.
    (a): the
    equilibrium state at $t=0$~\textsf{ps}. (b): The new equilibrium
    state at $t=1000$~\textsf{ps}, i.e., fully relaxed under the
    constant electric field 1~V/nm. The darker color
    indicates larger population on the Ramachandran histogram.}
  \label{fig:tmp4}
\end{figure}

@@

The Fig.~\ref{fig:tmp3}
shows that the $x$-component of the dipole moment of the alanine dipeptide
reaches 85\% of the maximum value in only 20~ps, which is comparable to
the warm-up time $t_{init}$. Then the dipole slowly relaxes to the 
maximum value in the following 400~ps. Notice that 20~ps is 25 times
smaller than the slowest time scale of the conformational relaxation.
We also consider the orientation of the alanine dipeptide as a function
of time. The geometric direction of the alanine is defined by the red vector in
Fig.~\ref{fig:tmp3}, which is the angle bisector of the two black vectors.
The black vectors are define to connect the $\alpha$-carbon and the carbons
on the methyl groups. We define the orientational order parameter
\begin{align}
  S_\theta = \langle 3\sin^2\theta - 2\rangle,
\end{align}
where $\theta$ is the angle between the orientation of the alanine dipeptide
and the direction of the electric field. The order parameter
indicate if the molecule is perpendicular to the electric field.
If the molecule is perfectly perpendicular to the electric field, then $S_\theta = 1$;
If the molecule has no directional
preference at all, then $S_\theta = 0$;
If the molecule is perfectly parallel to the electric field, then $S_\theta = -2$.
From $t=0$ to roughly 10~ps, the orientational order parameter fastly
decreases from 0 to -0.15, which means a weak alignment of the molecule to the external field.
Then from $t=0$ to 500~ps, the molecule slowy changes to the orientation
that is perpendicular to the electric field.

@@

>>>>

\begin{figure}
  \centering
  \raisebox{0.2\height}{\includegraphics[width=0.22\textwidth]{fig.ala/ext.mode1.010.Ex.01.00.t1000ps.recheck/c-2.eps}}
  \includegraphics[width=0.37\textwidth]{fig.ala/ext.mode1.010.Ex.01.00.t1000ps.recheck/fig-begin-1.eps}
  \includegraphics[width=0.37\textwidth]{fig.ala/ext.mode1.010.Ex.01.00.t1000ps.recheck/fig-end-1.eps}
  \caption{The population of metastable states of the alanine dipeptide.
    (a): the
    equilibrium state at $t=0$~\textsf{ps}. (b): The new equilibrium
    state at $t=1000$~\textsf{ps}, i.e., fully relaxed under the
    constant electric field 1~V/nm. The darker color
    indicates larger population on the Ramachandran histogram.}
  \label{fig:tmp4}
\end{figure}

@@

@@

The Fig.~\ref{fig:tmp3}
shows that the $x$-component of the dipole moment of the alanine dipeptide
reaches 85\% of the maximum value in only 20~ps, which is comparable to
the warm-up time $t_{init}$. Then the dipole slowly relaxes to the 
maximum value in the following 400~ps. Notice that 20~ps is 25 times
smaller than the slowest time scale of the conformational relaxation.
We calculate the averaged dipole moment of different metastable state
under the constant EF, and they are 6.8, 6.0, 5.1, 3.1 and 7.1~Debye
for $A_1$, $A_2$, $B_1$, $B_2$ and $C$, respectively.
Under a constant EF, the system will be driven towards those
conformations that with higher dipole moment, because the
energy of the system will be lowered by aligning the the dipole
to the external EF.
It is clear that
the $\alpha$-helix and left-handed $\alpha$-helix are of higher
dipole moments than $\beta$-sheet conformations, so the probability
of being in the $\alpha$-helix conformations were
observed being increased under the EF.  $A_1$ and $C$ are the most
probable conformations because their dipole moments are the highest
among all metastable conformation. Moreover, the $B_2$  vanishes, because
its dipole is significantly lower than other conformations.
% This also explains the two-time scale behavior of the molecular dipole
% moment.
From 0 to 20~ps, the probability of
lowest dipole conformation $B_2$ quickly
vanishes, so there is a sharp increment of the average molecular
dipole moment. Then until 500~ps, since the conformation slowly migrates from
$B_1$ to $A_1$ and $C$, and also since the dipole moment $A_1$ and $C$
is higher than $B_1$, the average molecular dipole moment slowly increases
to the maximum value.

We also consider the orientation of the alanine dipeptide as a function
of time.
The geometric direction of the alanine is defined by the red vector in
Fig.~\ref{fig:tmp3}, which is the angle bisector of the two black vectors.
The black vectors are define to connect the $\alpha$-carbon and the carbons
on the methyl groups. We define the orientational order parameter
\begin{align}
  S_\theta = \langle 3\sin^2\theta - 2\rangle,
\end{align}
where $\theta$ is the angle between the orientation of the alanine dipeptide
and the direction of the electric field. The order parameter
indicate if the molecule is perpendicular to the electric field.
If the molecule is perfectly perpendicular to the electric field, then $S_\theta = 1$;
If the molecule has no directional
preference at all, then $S_\theta = 0$;
If the molecule is perfectly parallel to the electric field, then $S_\theta = -2$.
From $t=0$ to roughly 10~ps, the orientational order parameter fastly
decreases from 0 to -0.15, which means a weak alignment of the molecule to the external field.
Then from $t=0$ to 500~ps, the molecule slowly changes to the orientation
that is perpendicular to the electric field.
